---
editor_options:
  chunk_output_type: console
output:
  html_document:
    fig_caption: yes
    figure_caption: yes
    highlight: tango
    number_sections: yes
    toc: yes
    toc_depth: 4
  pdf_document:
    toc: yes
    toc_depth: '4'
  word_document:
    toc: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Sampling rates for passive samplers exposed to a field-relevant peak of 42 organic pesticides

Verena C. Schreiner, Nikita Bakanov, Mira Kattwinkel, Sarah Könemann, Stefan Kunz, Etienne L.M. Vermeirssen & Ralf B. Schäfer

R script written by Verena C. Schreiner and Mira Kattwinkel
checked by Ralf B. Schäfer and Stefan Kunz
University of Koblenz-Landau  
Fortstrasse 7  
76829 Landau  
GERMANY  
Corresponding mail adress: schreiner-verena'@'uni-landau.de

## Tools / Packages
```{r, warning=FALSE,message=FALSE,error=FALSE}
library(simecol)      # algorithm used to calculate sampling rates of 5 and 7 day dataset
library(FME)          # algorithm used to calculate sampling rates of 7 day and combined dataset
library(data.table)   # data formatation
library(lubridate)    # formatation of dats + calculations of time lengths
library(glmnet)       # for LASSO approach
library(dplyr)        # data formatation
library(cvq2)         # calculate q2
```


## Load & Clean Data
```{r}
# project path, change to the path you saved the rawfiles
path <- getwd()
```

create folders in work directory where results and figures are saved later
```{r}
dir.create(file.path(path, "fits")) # creating a folder called "fits" (the graphs will be saved there)
dir.create(file.path(path, "TWA"))
```


### Load data
```{r}
# pesticide concentrations on SDB disks
Results_SDB = read.table(file.path(path,"Results_SDB.csv"), 
                            header=TRUE, sep = ";")
Results_SDB = melt(Results_SDB, id.vars = c("Compound"), value.name = "Concentration",
                   variable.name = "Code")
Results_SDB = as.data.table(Results_SDB)


# water concentrations of large injection system [ng/L] 
Results_EQUAN = read.table(file.path(path, "Results_EQuan.csv"), header = TRUE, sep = ";")
# transform to long version
Results_EQUAN = melt(Results_EQUAN, id.vars = c("Compound"), value.name = "Concentration", 
                      variable.name = "Code") 


# water concentrations via SPE [ng/L]
Results_SPE = read.table(file.path(path, "Results_SPE.csv"), header = TRUE, sep = ";")
Results_SPE = as.data.table(Results_SPE)
# transform to long version
Results_SPE = melt(Results_SPE, id.vars = c("Compound"), value.name = "Concentration", 
                      variable.name = "Code") 
# Warning because concentration of some compounds only complete values (int) and others with decimal (num)
# change codenames of column code
Results_SPE$Code = gsub("SPE.", "",Results_SPE$Code)
```

property data of compounds
```{r}
prop = read.table(file.path(path,"Info_compounds.csv"), 
                            header=TRUE, sep = ";", na.strings = c("NR", "NA"))
```

uptake rates from previous studies
```{r}
Rs_other_studies = read.table(file.path(path,"Rs_previous_publications.csv"), 
                            header=TRUE, sep = ";", na.strings = c("NR", "NA"))
```


### Clean data
codes SDBs
```{r}
# codes SDB disks
Code_SDB = read.table(file.path(path,"Code_SDB.csv"), 
                            header=TRUE, sep = ";")
Code_SDB = as.data.table(Code_SDB)

# delete rows with Blank (Charge Disk)
# Code_SDB <- Code_SDB[!Code_SDB$Charge_disks == "Blank",]
# CH12 und CH15 -> Blanks_VerII
Code_SDB[c(1:2), Charge_disks := c("Blank_VerII", "Blank_VerII")] # blanks equals Time where disks were add

# merge columsn with time information (with lubridate)
Code_SDB$Time = gsub("1899-12-31", "", Code_SDB$Time)

# add timepoint of blank
Code_SDB[c(1:2), Time := c("14:45:00", "14:45:00")]

Code_SDB = Code_SDB[, list(Code, Charge_disks, Vol_Part, 
                                           Vol_all, 
                           Time_retrieved = paste(dmy(Date), hms(Time)), comment)] # lubridate function

# change column Time_retrieved 
Code_SDB = Code_SDB[, list(Code, Charge_disks, Vol_Part, 
                                           Vol_all, 
                           Time_retrieved = strptime(Time_retrieved, "%Y-%Om-%Od %HH %MM %SS"), comment)]


# calculate time intervals of retrieval
# set of SDB fom start
VerI = Code_SDB[Charge_disks == "Ver I" | Charge_disks == "Blank",
                           list(Code, Charge_disks, Vol_Part, 
                                Vol_all, Time_retrieved,
                                Timespan_V1 = c(NA, 0, as.numeric(difftime(Time_retrieved[3:12], 
                                                                           Time_retrieved[2], units = "hours"))),
                                comment)]

# set of SDBs added later
VerII = Code_SDB[Charge_disks == "Ver II" | Charge_disks == "Blank_VerII", 
                            list(Code, Charge_disks, Vol_Part, 
                                 Vol_all, Time_retrieved,
                                 Timespan_V2 = c(NA, 0, as.numeric(difftime(Time_retrieved[3:8], 
                                                                            Time_retrieved[2], units = "hours"))),
                                 comment)]
# merge both dataframes
Code_SDB = rbind(VerI, VerII, fill = TRUE)
# exclude samples, where - since vial broke - not all pesticides could be retrieved
Code_SDB <- Code_SDB[! Code_SDB$Code == "CH6" & !Code_SDB$Code == "CH10",] 

# merge results SDB and codes
Results_SDB = merge(x = Results_SDB, y = Code_SDB, by = "Code")
# delete blanks (all lower than LOQ)
Results_SDB <- Results_SDB[!grep("^Blank", Results_SDB$Charge_disks),]
```

calculate absorbed mass [ng] 
0.1 mL or 100 uL is the injection volume
500 uL is the total volume of the condensed sample
```{r}
Results_SDB = Results_SDB[, list(Compound, Mass_ng = 
                                   (Concentration* 0.1 * 500/100 * (Vol_all/Vol_Part)), 
                                       Time_retrieved, Charge_disks, Timespan_V1, Timespan_V2,comment),
                                by = Code]

# sort for timepoint
Results_SDB = Results_SDB[order(Results_SDB$Time_retrieved),]
```


codes water samples
```{r}
Code_water = read.table(file.path(path,"Code_Water.csv"), 
                            header=TRUE, sep = ";")
Code_water = as.data.table(Code_water)
setnames(Code_water, c("code","extact.time"), c("Code","Timepoint"))

# formatting time column + sorting
Code_water$Timepoint = gsub("Uhr", "", Code_water$Timepoint)
Code_water = Code_water[, list(Code, Timepoint = strptime(Timepoint,"%d.%m.  %H:%M "), Time_after_Disk_1,
                                                 Time_after_Disk_2, comment)]


# calculate exact time point
Code_water = Code_water[, list(Code, Timepoint, Time_after_Disk_1 = 
                                                   c(0, cumsum(as.numeric(difftime(Timepoint[2:16], Timepoint[1:15],
                                                                                 units = "hours")))),
                                                 Time_after_Disk_2 = 
                                                   c(NA, NA, 0,cumsum(as.numeric(difftime(Timepoint[4:16],
                                                                                          Timepoint[3:15], 
                                                                                          units = "hours")))),
                                                 comment)]
# These two values are before SDB disks of set two (5-day dataset) were added
Code_water[1,4] <- c(-47.66667)
Code_water[2,4] <- c(-23.33333) 
```


combine both methods of determining water concentrations
```{r}
Results_Water =  rbind(Results_EQUAN, Results_SPE)
```

merge water concentrations with time points and sort
```{r}
Results_Water = merge(Results_Water, Code_water[,c(1:4)], by = "Code")

# order for timepoint
Results_Water = Results_Water[order(Results_Water$Timepoint),]
```


## calculate sampling rates
### prepare dataframes for 5- and 7-day approach
```{r}
Results_SDB = Results_SDB[order(Results_SDB$Compound),]
Results_Water = Results_Water[order(Results_Water$Compound),]
```


create new dataframe for the water concentrations with one time point
```{r}
Results_Water =  rbind(Results_EQUAN, Results_SPE)
Results_Water = merge(Results_Water, Code_water[,c(1,3)], by = "Code")

Results_Water = Results_Water[order(Results_Water$Time_after_Disk_1),]

Results_Water <- as.data.frame(Results_Water)

names(Results_Water)[3] <- c("Conc.water.D1") # concentrations of 7-day dataset
names(Results_Water)[4] <- c("Time.water")
# concentrations of 5-day dataset
Results_Water$Conc.water.D2 <- Results_Water$Conc.water.D1
Results_Water$Conc.water.D2 <- ifelse(Results_Water$Time.water < 30, 0, Results_Water$Conc.water.D2) 
Results_Water$Time.water <- (Results_Water$Time.water)/24 # use days instream of hours
Results_Water <- Results_Water[!is.na(Results_Water$Conc.water.D1),]
```


create dataframe for the passive sampler 
```{r}
Results_SDB <- as.data.frame(Results_SDB)
SDB_Results <- Results_SDB
# add nearly two days (exact time frame used) to 5-day approach to match with corresponding water concentrations
SDB_Results$Timespan_V2 <- SDB_Results$Timespan_V2 + 47.66667 
SDB_Results$Time.disks <- ifelse(is.na(SDB_Results$Timespan_V1), SDB_Results$Timespan_V2, SDB_Results$Timespan_V1)
SDB_Results$Charge_disks <- gsub("Ver II", "D2", SDB_Results$Charge_disks) # 5-day dataset
SDB_Results$Charge_disks <- gsub("Ver I", "D1", SDB_Results$Charge_disks) # 7-day dataset
SDB_Results <- SDB_Results[, c(2,3,5,9)]
names(SDB_Results)[2] <- c("mass_disk")
SDB_Results$Time.disks <- (SDB_Results$Time.disks)/24
SDB_Results <- SDB_Results[!is.na(SDB_Results$mass_disk),]
```


Remove outlier
```{r}
# when inspecting the different compounds, Epoxiconazole has one clear outlier
plot(SDB_Results[SDB_Results$Compound == "Epoxiconazol",]$Time.disks,
                 SDB_Results[SDB_Results$Compound == "Epoxiconazol",]$mass_disk)
# removing outlier
rm_Epox <- with(SDB_Results, (Compound == "Epoxiconazol") & (SDB_Results$mass_disk < 10))
SDB_Results <- SDB_Results[!rm_Epox, ]

```

save created dataframes
```{r}
write.csv(Results_Water, file = "Results_Water.csv")
write.csv(SDB_Results, file = "SDB_Results.csv")
```


### prepare dataframes
create list of analysed compounds
```{r}
Compounds = as.character(unique(Results_SDB$Compound))
# Compounds <- sort(Compounds, decreasing = FALSE) # sort Compoundnames (here not necessary, but maybe in future)
```

create sorption model of first order
with this the sampling rate can be calculated from the rate constants (will be calculated in the lower part)
the function for sorbent-water uptake under fixed water concentration (1 ug/L)
```{r}
sorb_model <- function(time, state, parms) { 
  with(as.list(c(state, parms)),{
    sorb <- state
    dsorb.dt <- + kws * 1 -  ksw * sorb
    list(c(dsorb.dt))
  })}
```


function to fit the rate constants (parameters)
```{r}
fun <- function(p, simObj, observations) {
  whichpar <- names(p)
  parms(simObj)[whichpar] <- p
  times(simObj) <- observations[,"time"]
  ysim <- out(sim(simObj))
  modCost(model = ysim, obs = observations, y = "val")
}
```


### combined loop for both single datasets and both datasets combined

create several empty data.frames, which will be filled in the loop
```{r}
k_values_t1_alg1 <- NULL
k_values_t1_alg2 <- NULL

Rs_t1_alg1 <- NULL
Rs_t1_alg2 <- NULL

k_values_t1 <- NULL
Rs_t1 <- NULL

k_values_t2 <- NULL
Rs_t2 <- NULL

k_values_comb <- NULL
Rs_comb <- NULL
deviance_comb <- NULL

# k_value_inst <- NULL
# Rs_inst <- NULL
# deviance_inst <- NULL

Rs_inst <- NULL
```


```{r, results="hide"}
for(i in 1:length(Compounds)) { # number of compounds 
y <- Compounds[i]

  # at this step calculations for compound y is conduced, will be replaes (in the loop) for each compound 
  Water_comb = Results_Water[Results_Water$Compound == y,] 
  # 7-day dataset
  Water_t1 = Results_Water[Results_Water$Compound == y, c(2,3,4)] 
  Water_t2 = Results_Water[Results_Water$Compound == y, c(2,5,4)] 
  # 5-day dataset: set the start of the water concentration to the stime where SDB disks are added
  Water_t2$Time.water <- Water_t2$Time.water - 1.9861111 
  Water_t2 <- Water_t2[!Water_t2$Time.water < 0,]
  
  SDB_comb = SDB_Results[SDB_Results$Compound == y ,]
  # 7-day dataset
  SDB_t1 = SDB_Results[SDB_Results$Compound == y & SDB_Results$Charge_disks == "D1",]
  SDB_t2 = SDB_Results[SDB_Results$Compound == y & SDB_Results$Charge_disks == "D2",]
  # 5-day dataset, set time when SDB disks are added to zero
  SDB_t2$Time.disks <- SDB_t2$Time.disks - 1.986083
  
  # mass of the sampler: 332 x 10-6 kg = 332 mg = 0.332 ng
  mass_samp <- 0.332
  

# combined dataset, algorithm 1
# given water concentrations 
# given values for kws and ksw (dummy variables)
mod6 <- new("odeModel",
            main = function (time, init, parms, ...) {
              with(as.list(c(init, parms, inputs)), {
                win1 <- approxTime1(inputs, time, rule=2)["w1"] # disks 7-day dataset
                win2 <- approxTime1(inputs, time, rule=2)["w2"] # disks 5-day dataset
                
                ds1 <- kws * win1 - ksw * s1
                ds2 <- kws * win2 - ksw * s2
                list(c(ds1, ds2))
              })
            },
            parms = c(kws=0.05, ksw=0.01),
            times =  c(from = 0, to=8, by=1/(24)),
            init = c(s1 = 0, s2 = 0),
            solver = "lsoda",
            initfunc = function(obj) {
              tt <- fromtoby(times(obj))
              inputs(obj) <- as.matrix(data.frame(
                time = Water_comb$Time.water, 
                w1 = Water_comb$Conc.water.D1, # water 7-day dataset
                w2 = Water_comb$Conc.water.D2 # water 5-day dataset
              ))
              obj
            }
)


### estimate rate constants
mod7 <- mod6

# observations, must be in this order!
obs_comb <- na.exclude(SDB_comb[,c("Charge_disks", "Time.disks",  "mass_disk")]) 
  
# same names as in model
obs_comb$Charge_disks[obs_comb$Charge_disks == "D1"] <- "s1"
obs_comb$Charge_disks[obs_comb$Charge_disks == "D2"] <- "s2"
colnames(obs_comb) <- c("name", "time", "val")
obs_comb <- obs_comb[order(obs_comb$time),] 


fit_comb <- modFit(f = fun, p = parms(mod7), lower = 0, simObj = mod7,  observations = obs_comb, method = "Nelder", control = list(trace=FALSE))
# print(fit_comb$par) # shows calculated rate constants


# add calculated rate constants to model (needed for graph below)
mod8 <- mod6
parms(mod8) <- fit_comb$par
res8 <- sim(mod8)


# save the rate constants to empty data frame
k_values_comb$kws[y] <- fit_comb$par[1] # kws
k_values_comb$ksw[y] <- fit_comb$par[2] # ksw


# t set to 5 days
Rs_comb[y] <- ((1 - exp(-fit_comb$par[2] * 5)) * fit_comb$par[1])/ (fit_comb$par[2] * 5)


# calculate model residual standard error
deviance_comb[y] <- summary(fit_comb)$sigma
# 


# instantaneous sampling rate
# given water concentrations 
# given values for Rs and k (dummy variables)
mod10 <- new("odeModel",
            main = function (time, init, parms, ...) {
              with(as.list(c(init, parms, inputs)), {
                win1 <- approxTime1(inputs, time, rule=2)["w1"] # disks 7-day dataset
                win2 <- approxTime1(inputs, time, rule=2)["w2"] # disks 5-day dataset
                
                ds1 <- Rs * (win1 -(s1/(mass_samp * k)))  # kws * win1 - ksw * s1
                ds2 <- Rs * (win2 -(s2/(mass_samp * k))) # kws * win2 - ksw * s2
                list(c(ds1, ds2))
              })
            },
            parms = c(Rs=0.1, k=0.01),
            times =  c(from = 0, to=8, by=1/(24)),
            init = c(s1 = 0, s2 = 0),
            solver = "lsoda",
            initfunc = function(obj) {
              tt <- fromtoby(times(obj))
              inputs(obj) <- as.matrix(data.frame(
                time = Water_comb$Time.water, 
                w1 = Water_comb$Conc.water.D1, # water 7-day dataset
                w2 = Water_comb$Conc.water.D2 # water 5-day dataset
              ))
              obj
            }
)


### estimate variables
mod11 <- mod10

# observations, must be in this order!
obs_comb <- na.exclude(SDB_comb[,c("Charge_disks", "Time.disks",  "mass_disk")]) 
  
# same names as in model
obs_comb$Charge_disks[obs_comb$Charge_disks == "D1"] <- "s1"
obs_comb$Charge_disks[obs_comb$Charge_disks == "D2"] <- "s2"
colnames(obs_comb) <- c("name", "time", "val")
obs_comb <- obs_comb[order(obs_comb$time),] 


fit_inst <- modFit(f = fun, p = parms(mod11), lower = 0, simObj = mod11,  observations = obs_comb, method = "Nelder", control = list(trace=F))
# print(fit_inst$par) # calculated parameter
# 
# 
# add calculated rate constants to model (needed for graph below)
mod12 <- mod10
parms(mod12) <- fit_inst$par
res8 <- sim(mod12)
# 


Rs_inst$Rs_initial[y] <- fit_inst$par[1]
Rs_inst$k[y] <- fit_inst$par[2]

# calculate model residual standard error
Rs_inst$deviance[y] <- summary(fit_inst)$sigma


### Timepoint 1  - 7-day dataset
### Algorithm 1
# given water concentrations 
# given (dummy) values for kws and ksw
mod1 <- new("odeModel",
              main = function (time, init, parms, ...) {
               with(as.list(c(init, parms, inputs)), {
                win <- approxTime1(inputs, time, rule=2)["w"]  
                # with this, the water concontration of every point in time is approximated
                ds <- kws * win - ksw * s
                list(c(ds))
               })
              },
              parms = c(kws=0.05, ksw=0.01),  # adding approximate parameters for the fits
            ### possibility to change analysed time frames
            ### e.g. times = c(from=0, to=16, by=1/(24 * 2)), for half hour
            ### will take longer to run
              times = c(from=0, to=8, by=1/(24)),# c(from=0, to=170, by=1/(24))
            # everything will be modelled for each hour a day
              init = c(s = 0), # initial concentration on the sorbent (s) is 0
              solver = "lsoda", # the used algorith to solve the differential equation
            # lsoda is used as a algorithm due to its flexibility, 
            # which is necessary because of the large steps in the water concentration
              initfunc = function(obj) {
                tt <- fromtoby(times(obj))
                inputs(obj) <- as.matrix(data.frame(
                  time = Water_t1$Time.water, 
                  w = Water_t1$Conc.water.D1
                  # building a matrix with the measured water concentrations, will be used in upper part of model
                ))
                obj
              }
)


### estimate rate constants ###
mod2 <- mod1

# observations, must be in this order!
obs_t1 <- na.exclude(SDB_t1[,c("Charge_disks", "Time.disks",  "mass_disk")]) 
  
# same names as in model
obs_t1$Charge_disks <- "s"
colnames(obs_t1) <- c("name", "time", "val")
obs_t1 <- obs_t1[order(obs_t1$time),] # time needs to be in the correct order, otherwise an error will occur

fit_t1_alg1 <- modFit(f = fun, p = parms(mod2),lower = 0, simObj = mod2,  observations = obs_t1, method = "Nelder", control = list(trace=FALSE))
# print(fit_t1_alg1$par) # estimated rate constants

# the calculated rate constants are added to the model
# used to later on prepare graph
mod3_alg1 <- mod1
parms(mod3_alg1) <- fit_t1_alg1$par
res3_alg1 <- sim(mod3_alg1)


# save the rate constants to data.frame
k_values_t1_alg1$kws[y] <- fit_t1_alg1$par[1] # kws
k_values_t1_alg1$ksw[y] <- fit_t1_alg1$par[2] # ksw

# t set to 5 days
Rs_t1_alg1[y] <- ((1 - exp(-fit_t1_alg1$par[2] * 5)) * fit_t1_alg1$par[1])/ (fit_t1_alg1$par[2] * 5)



### Timepoint 1  - 7-day dataset
### Algorithm 2
obstime_t1 <- SDB_t1$Time.disks             # measured times
i_t1 <- which(!is.na(obstime_t1))            # delete lines with NA
obstime_t1 <- obstime_t1[i_t1]
yobs_t1 <- data.frame(s = SDB_t1$mass_disk[i_t1])  # add concentration values where time is also available

### fit  with range of starting values for kws and ksw values
### you can try different ranges and starting values for other substances/ substance classes
mod3_alg2 <- mod1
lower <- c(kws = 0, ksw = 0)   # lower range of tested values
upper <- c(kws = 10, ksw = 10) # upper range of tested values
# this range in added to reduce the time the algorithm needs to calculate, range could be extended or reduced
parms(mod3_alg2) <- c(kws = 1, ksw = 0.1) # adding dummy variables (idealy in the range you would expect them)
fit_t1_alg2 <- fitOdeModel(mod3_alg2, lower = lower, upper = upper, obstime = obstime_t1,
                    yobs = yobs_t1, control=list(trace = FALSE))
# using fitODEModel the rate constants are fitted with the observations to calculate the most reasonable ones


# save the rate constants
k_values_t1_alg2$kws[y] <- fit_t1_alg2@par[1] # kws
k_values_t1_alg2$ksw[y] <- fit_t1_alg2@par[2] # ksw

# t set to 5 days
Rs_t1_alg2[y] <- ((1 - exp(-fit_t1_alg2@par[2] * 5)) * fit_t1_alg2@par[1])/ (fit_t1_alg2@par[2] * 5)




### Timepoint 2  - 5-day dataset
### algorithm 2
mod4 <- new("odeModel",
              main = function (time, init, parms, ...) {
               with(as.list(c(init, parms, inputs)), {
                win <- approxTime1(inputs, time, rule=2)["w"]  
                # With this, the water concontration of every point in time is approximated
                ds <- kws * win - ksw * s
                list(c(ds))
               })
              },
              parms = c(kws=1, ksw=0.1),  # adding approximate paramaters for the fits
              times = c(from=0, to=8, by=1/(24)), 
            # everything will be modelled for each hour a day
              init = c(s = 0), # initial concentration on the sorbent (s) is 0
              solver = "lsoda", # the used algorith to solve the differential equation
            # lsoda is used as a algorithm due to its flexibility, 
            # which is necessary because of the large steps in the water concentration
              initfunc = function(obj) {
                tt <- fromtoby(times(obj))
                inputs(obj) <- as.matrix(data.frame(
                  time = Water_t2$Time.water, 
                  w = Water_t2$Conc.water.D2
                  # building a matrix with the measured water concentrations, will be used in upper part of model
                ))
                obj
              }
)

### estimate rate constants
mod5 <- mod4

# observations of sorbent for 5-day dataset
obstime_t2 <- SDB_t2$Time.disks             # measured times
i_t2 <- which(!is.na(obstime_t2))            # delete lines with NA
obstime_t2 <- obstime_t2[i_t2]
yobs_t2 <- data.frame(s = SDB_t2$mass_disk[i_t2])  # add concentration values where time is also available


lower <- c(kws = 0, ksw = 0)   # lower range of rate constants
upper <- c(kws = 10, ksw = 10) # upper range of rate constants
# this range in added to reduce the time the algorithm needs to calculate, range could be extended or reduced
parms(mod5) <- c(kws = 1, ksw = 0.1) # adding dummy variables (idealy in the range you would expect them)
fit_t2 <- fitOdeModel(mod5, lower = lower, upper=upper, obstime = obstime_t2,
                   yobs = yobs_t2, control=list(trace = FALSE))
# using fitODEModel the k-values are fitted with the observations to calculate the most reasonable ones

# save the rate constants to empty data frame
k_values_t2$kws[y] <- fit_t2@par[1] # kws
k_values_t2$ksw[y] <- fit_t2@par[2] # ksw

# t set to 5 days
Rs_t2[y] <- ((1 - exp(-fit_t2@par[2] * 5)) * fit_t2@par[1])/ (fit_t2@par[2] * 5)


# plotting + saving each of the fits
file = y # for compound y
file = file.path(path,"fits", paste(file,"png", sep=".")) # create filename for each figure, will be saves in newly created folfer "fits"
png(filename=file, width=7, height=6, units="in", res=1000)
par(mar=c(5.1, 6.9, 2, 1), mgp = c(3.5,1,0))

# Prepare one figure with all sample points + fit
# create empty plot
plot(res3_alg1, ylim = c(0, (12/10)*max(Water_t1$Conc.water.D1)), main = paste(y), lwd = 2, xlab = c("time [days]"),
     ylab = expression(atop("water concentration [ng L"^"-1"* "]", "compound mass on disk [ng]")), 
     xlim = c(0,8), las = 1, cex.lab = 1.5, cex.axis = 1.5, cex.main = 1.5, xaxs="i", yaxs="i", type = "n") 


# add 7-day data
points(SDB_t1$Time.disks, SDB_t1$mass_disk, pch = 16) # the points of the SDB disk concentration are added 

# add 5-day data
points(SDB_t2$Time.disks + (47.666/24)
       , SDB_t2$mass_disk, col="red", pch = 16) # the points of the SDB disk concentration are added 

# add results of fit
lines(res8@out , col="darkgreen", lwd = 2) 

# add water concentrations with loess smoother
loess_water <- loess(Conc.water.D1  ~ Time.water, data = Water_t1, span = 0.3)
lines(Water_t1$Time.water, predict(loess_water), col = "blue", lwd = 2)
points(Water_t1$Time.water, Water_t1$Conc.water.D1, col = "blue", pch = 16) # the water concentration is added as observed


# Figures used for Figure S2 and Figure of diazinon is displayed in the manuscript as Figure 1
dev.off()


}
```

### saving results into one dataframe

rate constants
```{r}
# 7-day dataset alg. 1
k_values_t1_alg1 <- as.data.frame.list(k_values_t1_alg1)
k_values_t1_alg1$Compound <- rownames(k_values_t1_alg1)

setnames(k_values_t1_alg1, c("ksw","kws"), c("ksw_t1_alg1", "kws_t1_alg1"))

# 7-day dataset alg. 2
k_values_t1_alg2 <- as.data.frame.list(k_values_t1_alg2)
k_values_t1_alg2$Compound <- rownames(k_values_t1_alg2)

setnames(k_values_t1_alg2, c("ksw","kws"), c("ksw_t1_alg2", "kws_t1_alg2"))

# 5-day dataset
k_values_t2 <- as.data.frame.list(k_values_t2)
k_values_t2$Compound <- rownames(k_values_t2) 

setnames(k_values_t2, c("ksw","kws"), c("ksw_t2", "kws_t2"))

# combined dataset
k_values_comb <- as.data.frame.list(k_values_comb)
k_values_comb$Compound <- rownames(k_values_comb)

deviance_comb <- as.data.frame.list(deviance_comb)
deviance_comb <- t(deviance_comb)
k_values_comb <- cbind(k_values_comb, deviance_comb)
setnames(k_values_comb, c("ksw","kws"), c("ksw_comb", "kws_comb"))

# combine all dataframes
k_values <- merge( k_values_t1_alg1, k_values_t1_alg2, by=c("Compound")) 
k_values <- merge(k_values, k_values_t2, by=c("Compound")) 
k_values <- merge(k_values, k_values_comb, by=c("Compound")) 
```


sampling rates
```{r}
# 7-day dataset alg. 1
Rs_t1_alg1 <- as.data.frame(t(as.data.frame.list(Rs_t1_alg1)))
Rs_t1_alg1$Compounds <- rownames(Rs_t1_alg1)
names(Rs_t1_alg1)[1] <- c("Rs_t1_alg1")
rownames(Rs_t1_alg1) <- c() # get rid of weird row names

# 7-day dataset alg. 2
Rs_t1_alg2 <- as.data.frame(t(as.data.frame.list(Rs_t1_alg2)))
Rs_t1_alg2$Compounds <- rownames(Rs_t1_alg2)
names(Rs_t1_alg2)[1] <- c("Rs_t1_alg2")
rownames(Rs_t1_alg2) <- c() # get rid of weird row names

# 5-day dataset
Rs_t2 <- as.data.frame(t(as.data.frame.list(Rs_t2)))
Rs_t2$Compounds <- rownames(Rs_t2)
names(Rs_t2)[1] <- c("Rs_t2")
rownames(Rs_t2) <- c() # get rid of weird row names

# combined dataset
Rs_comb <- as.data.frame(t(as.data.frame.list(Rs_comb)))
Rs_comb$Compounds <- rownames(Rs_comb)
names(Rs_comb)[1] <- c("Rs_comb")
rownames(Rs_comb) <- c() # get rid of weird row names

# instantaneous sampling rate
Rs_inst_back <- Rs_inst
Rs_inst <- as.data.frame(as.data.frame.list(Rs_inst))
Rs_inst$Compound <- rownames(Rs_inst)
rownames(Rs_inst) <- c() # get rid of weird row names

# combine dataframes
Rs <- merge(Rs_t1_alg1, Rs_t1_alg2, by = c("Compounds"))
Rs <- merge(Rs, Rs_t2, by = c("Compounds"))
Rs <- merge(Rs, Rs_comb, by = c("Compounds"))

# correct compound names
setnames(Rs, c("Compounds"), c("Compound"))
Rs$Compound <- gsub("X2.4.D", "2-4-D", Rs$Compound)
Rs$Compound <- gsub("X2.n.Octyl.4.isothiazolin.3.on..OIT.", "2-n-Octyl-4-isothiazolin-3-on-(OIT)", Rs$Compound)
Rs$Compound <- gsub("Benthiavalicarb.isopropyl", "Benthiavalicarb-isopropyl", Rs$Compound)
Rs$Compound <- gsub("Piperonyl.butoxide", "Piperonyl-butoxide", Rs$Compound)

# add initial sampling rate
Rs <- merge(Rs, Rs_inst, by = c("Compound"))
```

save dataframe
```{r}
results_all <- merge(Rs, k_values, by = c("Compound"))
write.csv(results_all, file = "results_all2.csv", row.names = FALSE)
```

### Figure S1: Relationship between time-dependent 5-day sampling rate and initial sampling rate
```{r}
Rs_vergl <- merge(Rs_inst[,c(1,4)], results_all[,c(1,5)], by = c("Compound"))
```

```{r}
png(filename="Rs_vgl_initial.png", width=6, height=5, units="in", res = 1000) 
par(mar=c(4.1, 4.9, 2.1, 2.1))
plot(Rs_vergl$Rs_initial, Rs_vergl$Rs_comb, pch = 19, xlab = expression(paste("R"["s_t0"]*" [L d "^"-1"*"]")), ylab = expression(paste("R"["s_5d"]*" [L d "^"-1"*"]")))
abline(0,1, col = "black", lwd = 2) # identity line, 1:1 line
dev.off()
```


### Figure S6: relationship sampling rates timepoint 1 two different algorithms
```{r}
png(filename="Rs_7_day_dataset.png", width=6, height=5, units="in", res = 1000) 
par(mar=c(4.1, 4.9, 2.1, 2.1))
Rs_both_t1_2alg <- lm(Rs_t1_alg1 ~ Rs_t1_alg2, data = Rs) 
plot(Rs$Rs_t1_alg2, Rs$Rs_t1_alg1, pch=16, ylim=c(0,1), xlim=c(0,1), ylab = expression(paste("Rs using algorithm 1 [L d  "^"-1"*"]")), xlab = expression(paste("Rs using algorithm 2 [L d "^"-1"*"]")), xaxs="i", yaxs="i")

abline(Rs_both_t1_2alg)
text(0.1, 0.9, "r = 0.998 \n p < 0.001 \n n = 42")
summary(Rs_both_t1_2alg) # rÂ² = 0.9975 
cor(Rs$Rs_t1_alg1, Rs$Rs_t1_alg2)
dev.off()
```


## LASSO approach - estimate sampling rates with compound properties

### calculate logDow from logKow and from pKa - source ppdb
To know the water solubility for neutral fraction
calculation according to Carmichael, 2014
```{r}
prop$logDow <- ifelse(is.na(prop$comment), prop$logkow, 
                      ifelse(prop$comment == "acid", prop$logkow - log10(1 + 10^(8 - prop$pka)), 
                      ifelse (prop$comment == "base", prop$logkow - log10(1 + 10^(prop$pka - 8)), NA)))# 
# pH of 8 which was the pH in our channel system
# for acids: pH - pKa
# for bases: pKa - pH
```

### combine dataframes

combine sampling rates with compound properties
use sampling rates of combined dataset
```{r}
Rs_prop <- merge(Rs[,c(1,5)], prop, by = "Compound")
```

divide into pesticide groups
```{r}
Rs_prop_H <- Rs_prop[Rs_prop$type == "H",]
Rs_prop_I <- Rs_prop[Rs_prop$type == "I",]
Rs_prop_F <- Rs_prop[Rs_prop$type == "F",]
Rs_prop_S <- Rs_prop[Rs_prop$type == "S",]

nrow(Rs_prop) == nrow(Rs_prop_H) + nrow(Rs_prop_I) + nrow(Rs_prop_F) + nrow(Rs_prop_S) # should be true
```


### conduct 1. LASSO for explanatory variables/ compound properties available for all compounds
use time-dependent sampling rate
check variable
```{r}
hist(Rs_prop$ws)
```
--> not normally distributed --> log transformation
```{r}
Rs_prop$log_ws <- log10(Rs_prop$ws)
hist(Rs_prop$log_ws)
```
--> normally distributed



```{r}
exp <- select(Rs_prop, logkow, mass, log_ws, logDow)

mt_ex <- as.matrix(exp) # --> X
res_iv <- as.matrix(select(Rs_prop, Rs_comb)) # --> Y

# LASSO WITH ALPHA = 1
cv1 <- cv.glmnet(mt_ex, res_iv, family = "gaussian", nfold = 4, type.measure = "deviance", paralle = TRUE, alpha = 1)
md1 <- glmnet(mt_ex, res_iv, family = "gaussian", lambda = cv1$lambda.1se, alpha = 1)
coef(md1)
# No variable was selected
```

### conduct 1. LASSO for explanatory variables/ compound properties only available for a subset of compounds
use time-dependent sampling rate
```{r}
# make subset of compounds
sub <- select(Rs_prop, Compound, Rs_comb, kfoc, pka)
sub <- sub[!is.na(sub$pka),]
sub <- sub[!is.na(sub$kfoc),]

exp_sub <- select(sub, kfoc, pka)

mt_ex_sub <- as.matrix(exp_sub) # --> X
res_iv_sub <- as.matrix(select(sub, Rs_comb)) # --> Y

# LASSO WITH ALPHA = 1
cv1_sub <- cv.glmnet(mt_ex_sub, res_iv_sub, family = "gaussian", nfold = 4, type.measure = "deviance", paralle = TRUE, alpha = 1)
md1_sub <- glmnet(mt_ex_sub, res_iv_sub, family = "gaussian", lambda = cv1$lambda.1se, alpha = 1)
coef(md1_sub)
# no variable selected
```

### conduct 2. LASSO for pesticides which remain in the linear uptake
exclude 17 compounds, which approach equilibrium
```{r}
comp_equ2 <- c("2-4-D", "2-n-Octyl-4-isothiazolin-3-on-(OIT)", "Acetamiprid", "Boscalid", "Carbendazim", "Cyproconazol", "Dimethoat", "Dimethomorph", "Imidacloprid", "Iprovalicarb", "MCPA", "Nicosulfuron", "Propamocarb", "Prosulfocarb", "Thiabendazol", "Thiacloprid", "Trifloxystrobin")
```

```{r}
Rs_prop <- as.data.table(Rs_prop)
Rs_prop_equ2 <- Rs_prop[ ! Rs_prop$Compound %in% comp_equ2, ]
```

check variable
```{r}
hist(Rs_prop_equ2$ws)
```
--> not normally distributed --> log transformation
```{r}
Rs_prop_equ2$log_ws <- log10(Rs_prop_equ2$ws)
hist(Rs_prop_equ2$log_ws)
```
--> normally distributed


```{r}
exp_equ2 <- select(Rs_prop_equ2, logkow, mass, log_ws, logDow)

mt_ex_equ2 <- as.matrix(exp_equ2) # --> X
res_iv_equ2 <- as.matrix(select(Rs_prop_equ2, Rs_comb)) # --> Y

# LASSO WITH ALPHA = 1
cv1_equ2 <- cv.glmnet(mt_ex_equ2, res_iv_equ2, family = "gaussian", nfold = 4, type.measure = "deviance", paralle = TRUE, alpha = 1)
md1_equ2 <- glmnet(mt_ex_equ2, res_iv_equ2, family = "gaussian", lambda = cv1_equ2$lambda.1se, alpha = 1)
coef(md1_equ2)
# no variable was selected
# very low numbers equal 0!
```


### conduct 2. LASSO for explanatory variables/ compound properties only available for a subset of compounds
only for compounds which remain in the linear uptake
exclude 17 compounds, which approach equilibrium

```{r}
# make subset of compounds
sub_equ2 <- select(Rs_prop_equ2, Compound, Rs_comb, kfoc, pka)
sub_equ2 <- sub_equ2[!is.na(sub_equ2$pka),]
sub_equ2 <- sub_equ2[!is.na(sub_equ2$kfoc),]

exp_sub_equ2 <- select(sub_equ2, kfoc, pka)

mt_ex_sub_equ2 <- as.matrix(exp_sub_equ2) # --> X
res_iv_sub_equ2 <- as.matrix(select(sub_equ2, Rs_comb)) # --> Y

# LASSO WITH ALPHA = 1
cv1_sub_equ2 <- cv.glmnet(mt_ex_sub_equ2, res_iv_sub_equ2, family = "gaussian", nfold = 4, type.measure = "deviance", paralle = TRUE, alpha = 1)
md1_sub_equ2 <- glmnet(mt_ex_sub_equ2, res_iv_sub_equ2, family = "gaussian", lambda = cv1_sub_equ2$lambda.1se, alpha = 1)
coef(md1_sub_equ2)
# no variable selected
# very low numbers equal 0!
```



### conduct 3. LASSO for pesticides for initial sampling rate

```{r}
Rs_prop2 <- Rs_prop
Rs_prop2 <- merge(Rs_prop2, Rs_inst[,c(1,4)], by = c("Compound"))
```

check variable
```{r}
hist(Rs_prop2$ws)
```
--> not normally distributed --> log transformation
```{r}
Rs_prop2$log_ws <- log10(Rs_prop2$ws)
hist(Rs_prop2$log_ws)
```
--> normally distributed


```{r}
exp_equ3 <- select(Rs_prop2, logkow, mass, log_ws, logDow)

mt_ex_equ3 <- as.matrix(exp_equ3) # --> X
res_iv_equ3 <- as.matrix(select(Rs_prop2, Rs_initial)) # --> Y

# LASSO WITH ALPHA = 1
cv1_equ3 <- cv.glmnet(mt_ex_equ3, res_iv_equ3, family = "gaussian", nfold = 4, type.measure = "deviance", paralle = TRUE, alpha = 1)
md1_equ3 <- glmnet(mt_ex_equ3, res_iv_equ3, family = "gaussian", lambda = cv1_equ3$lambda.1se, alpha = 1)
coef(md1_equ3)
# no variable was selected!
```


### conduct 3. LASSO for explanatory variables/ compound properties only available for a subset of compounds
for initial sampling rate

```{r}
# make subset of compounds
sub_equ3 <- select(Rs_prop2, Compound, Rs_initial, kfoc, pka)
sub_equ3 <- sub_equ3[!is.na(sub_equ3$pka),]
sub_equ3 <- sub_equ3[!is.na(sub_equ3$kfoc),]

exp_sub_equ3 <- select(sub_equ3, kfoc, pka)

mt_ex_sub_equ3 <- as.matrix(exp_sub_equ3) # --> X
sub_equ3 <- as.data.table(sub_equ3)
res_iv_sub_equ3 <- as.matrix(select(sub_equ3, Rs_initial)) # --> Y

# LASSO WITH ALPHA = 1
cv1_sub_equ3 <- cv.glmnet(mt_ex_sub_equ3, res_iv_sub_equ3, family = "gaussian", nfold = 4, type.measure = "deviance", paralle = TRUE, alpha = 1)
md1_sub_equ3 <- glmnet(mt_ex_sub_equ3, res_iv_sub_equ3, family = "gaussian", lambda = cv1_sub_equ3$lambda.1se, alpha = 1)
coef(md1_sub_equ3)
# no variable selected
```


### Figure S5a: relationship of log Kow to sampling rate
octanol-water partition coefficient - log10
check variable
```{r}
hist(Rs_prop$logkow)
```
--> more or less normally distributed

all compounds
```{r}
Kow_all <- lm(Rs_comb ~ logkow, data = Rs_prop) 
png(filename="Rs_to_Kow.png", width=6, height=5, units="in", res = 1000) 
par(mar=c(4.1, 4.9, 2.1, 2.1))
plot(Rs_prop$logkow, Rs_prop$Rs_comb, type = "n", pch=16, ylim=c(0,1.1), xlim=c(-0.3, 5.2), cex.axis = 1.5, cex.lab = 1.5, xlab = expression(paste("log K"[ow])), ylab = expression(paste("Rs [L d "^"-1"*"]")), las = 1, xaxs="i", yaxs="i")
abline(Kow_all)
points(Rs_prop_H$logkow, Rs_prop_H$Rs_comb, pch=1, cex = 2) # herbicides
points(Rs_prop_I$logkow, Rs_prop_I$Rs_comb, pch=16, col = "black", cex = 2) #insecticides
points(Rs_prop_F$logkow, Rs_prop_F$Rs_comb, pch=16, col = "gray50", cex = 2) # fungicides
points(Rs_prop_S$logkow, Rs_prop_S$Rs_comb, pch=10, col = "black", cex = 2) # synergist

summary(Kow_all)

text(0.2, 1.0, "R² = 0.25 \n p < 0.001 \n n = 42")
dev.off()
```


### Figure S5b: relationship of log Dow (ppdb) to sampling rate
distribution coefficient
```{r}
hist(Rs_prop$logDow)
```
--> more or less normally distributed

all compounds
```{r}
Dow_all <- lm(Rs_comb ~ logDow, data = Rs_prop) 
png(filename="Rs_to_Dow.png", width=6, height=5, units="in", res = 1000) 
par(mar=c(4.1, 4.9, 2.1, 2.1))
plot(Rs_prop$logDow, Rs_prop$Rs_comb, type = "n", pch=16, ylim=c(0,1.1), xlim=c(-6.9, 5), cex.axis = 1.5, cex.lab = 1.5, xlab = expression(paste("log D"[ow])), ylab = expression(paste("Rs [L d "^"-1"*"]")), las = 1, xaxs="i", yaxs="i")
abline(Dow_all)
points(Rs_prop_H$logDow, Rs_prop_H$Rs_comb, pch=1, cex = 2) # herbicides
points(Rs_prop_I$logDow, Rs_prop_I$Rs_comb, pch=16, col = "black", cex = 2) #insecticides
points(Rs_prop_F$logDow, Rs_prop_F$Rs_comb, pch=16, col = "gray50", cex = 2) # fungicides
points(Rs_prop_S$logDow, Rs_prop_S$Rs_comb, pch=10, col = "black", cex = 2) # synergist

summary(Dow_all) 

text(-5.7, 0.95, "R² = 0.13 \n p = 0.019 \n n = 42")
dev.off()
```

### Figure S5c: relationship of log Dow (ChemAxon) to sampling rate
```{r}
hist(Rs_prop$logDow_chem)
```
--> more or less normally distributed

```{r}
Dow_all_chem <- lm(Rs_comb ~ logDow_chem, data = Rs_prop) 
png(filename="Rs_to_Dow_chem.png", width=6, height=5, units="in", res = 1000) 
par(mar=c(4.1, 4.9, 2.1, 2.1))
plot(Rs_prop$logDow_chem, Rs_prop$Rs_comb, type = "n", pch=16, ylim=c(0,1), xlim=c(-1.5, 5.3), cex.axis = 1.5, cex.lab = 1.5, xlab = expression(paste("log D"[ow])), ylab = expression(paste("Rs [L d "^"-1"*"]")), las = 1, xaxs="i", yaxs="i")
abline(Dow_all_chem)
points(Rs_prop_H$logDow_chem, Rs_prop_H$Rs_comb, pch=1, cex = 2) # herbicides
points(Rs_prop_I$logDow_chem, Rs_prop_I$Rs_comb, pch=16, col = "black", cex = 2) #insecticides
points(Rs_prop_F$logDow_chem, Rs_prop_F$Rs_comb, pch=16, col = "gray50", cex = 2) # fungicides
points(Rs_prop_S$logDow_chem, Rs_prop_S$Rs_comb, pch=10, col = "black", cex = 2) # synergist
summary(Dow_all_chem) 

text(-0.85, 0.9, "R² = 0.35 \n p < 0.001 \n n = 42")
dev.off()
```


## calculate TWA concentrations from sampling rates and mass absorbed to disks with longest exposure durations
### 5-day dataset
```{r}
Results_SDB_5_day <- Results_SDB[Results_SDB$Code == "CH1" | Results_SDB$Code == "CH5",] 
# SDB disks CH1 and CH5 are the ones which were exposed for 5 days
# calculate mean of both replicates
Results_SDB_5_day <- aggregate(Results_SDB_5_day$Mass_ng, list(Results_SDB_5_day$Compound, Results_SDB_5_day$Timespan_V2), mean)
names(Results_SDB_5_day)[1] <- c("Compound")
names(Results_SDB_5_day)[2] <- c("Timespan")
names(Results_SDB_5_day)[3] <- c("Mass_ng")
Results_SDB_5_day$Timespan <- Results_SDB_5_day$Timespan/24 # change to days
```

calculate different TWAs
```{r}
TWA_SDB_5 <- merge(Results_SDB_5_day, Rs[,c(1,5)], by = c("Compound"))
TWA_SDB_5$TWA_5d <- TWA_SDB_5$Mass_ng / (TWA_SDB_5$Rs_comb * TWA_SDB_5$Timespan) 
# TWA calculated over acutal exposed time frame
TWA_SDB_5$TWA_2d_5da <- TWA_SDB_5$Mass_ng / (TWA_SDB_5$Rs_comb * 2) 
# TWA calculated for SDBs exposed for 5 days with exposure time of 2 days (equation 4 in publication)
```


### 7-day dataset
```{r}
Results_SDB_7_day <- Results_SDB[Results_SDB$Code == "CH18",] 
# sample CH18 was exposed for 7 days, 2nd replicate was lost
Results_SDB_7_day <- aggregate(Results_SDB_7_day$Mass_ng, list(Results_SDB_7_day$Compound, Results_SDB_7_day$Timespan_V1), mean)
names(Results_SDB_7_day)[1] <- c("Compound")
names(Results_SDB_7_day)[2] <- c("Timespan")
names(Results_SDB_7_day)[3] <- c("Mass_ng")
Results_SDB_7_day$Timespan <- Results_SDB_7_day$Timespan/24
```

calculate different TWAs
```{r}
TWA_SDB_7 <- merge(Results_SDB_7_day, Rs[,c(1,5)], by = c("Compound"))
TWA_SDB_7$TWA_7d <- TWA_SDB_7$Mass_ng / (TWA_SDB_7$Rs_comb * TWA_SDB_7$Timespan) 
# TWA calculated over acutal exposed time frame
TWA_SDB_7$TWA_2d_7da <- TWA_SDB_7$Mass_ng / (TWA_SDB_7$Rs_comb * 2) 
# TWA calculated for SDBs exposed for 7 days with exposure time of 2 days (equation 4 in publication)
```

### combine both dataframes
```{r}
TWA_SDB <- cbind(TWA_SDB_5, TWA_SDB_7)
# only keep relevant columns
TWA_SDB <- TWA_SDB[c("Compound", "TWA_5d", "TWA_7d", "TWA_2d_5da", "TWA_2d_7da")]
```


### compare different TWAs to highest (peak) water concentrations
```{r}
Results_Water_max <-  aggregate(Conc.water.D1 ~ Compound, Results_Water, FUN=max)
```

```{r}
compare_TWA_conc_max <- merge(Results_Water_max, TWA_SDB, by = c("Compound"))

compare_TWA_conc_max$comp_7d <- (compare_TWA_conc_max$TWA_7d / compare_TWA_conc_max$Conc.water.D1)*100
compare_TWA_conc_max$comp_5d <- (compare_TWA_conc_max$TWA_5d / compare_TWA_conc_max$Conc.water.D1)*100
compare_TWA_conc_max$comp_2d_7da <- (compare_TWA_conc_max$TWA_2d_7da / compare_TWA_conc_max$Conc.water.D1)*100
compare_TWA_conc_max$comp_2d_5da <- (compare_TWA_conc_max$TWA_2d_5da / compare_TWA_conc_max$Conc.water.D1)*100

# ratio to maximum (peak) water concentration
# 7 days exposed
mean(compare_TWA_conc_max$comp_7d) # 14.5%
sd(compare_TWA_conc_max$comp_7d) # 4.1
# 5 days exposed
mean(compare_TWA_conc_max$comp_5d) # 21.4%
sd(compare_TWA_conc_max$comp_5d) # 5.0

# 7 days exposed use timeframe of 2 days
mean(compare_TWA_conc_max$comp_2d_7da) # 51.0%
sd(compare_TWA_conc_max$comp_2d_7da) # 14.3

# 5 days exposed use timeframe of 2 days
mean(compare_TWA_conc_max$comp_2d_5da) # 53.8%
sd(compare_TWA_conc_max$comp_2d_5da) # 12.6
```


### prepare Figure 3 
```{r}
  # only use Diazinon
  Water_t1 = Results_Water[Results_Water$Compound == "Diazinon",] 
  TWA = TWA_SDB[TWA_SDB$Compound == "Diazinon",]

png(filename="Figure_3.png", width=7, height=6, units="in", res=1000)
par(mar=c(4.1, 4.9, 2.1, 2.1))
# create empty plot
plot(Water_t1$Time.water, Water_t1$Conc.water.D1, type ="n", pch = 16, ylim = c(0, (11/10)*max(Water_t1$Conc.water.D1)),
     xlim = c(0,7), main = "Diazinon", lwd = 2, ylab = expression(paste("concentration [ng L"^"-1"*"]")), xlab = c("time [days]"), las = 1, cex.axis = 1.2, cex.lab = 1.2, xaxs="i", yaxs="i")


# Water concentration
loess_water <- loess(Conc.water.D1  ~ Time.water, data = Water_t1, span = 0.3)
lines(Water_t1$Time.water, predict(loess_water), col = "blue", lwd = 2)
points(Water_t1$Time.water, Water_t1$Conc.water.D1, col = "blue", pch = 16)

abline(h = TWA$TWA_7d, col = "darkgreen", lwd = 2.5) # add TWA SDB disk exposed 7 days
lines(c(2,7.5),c(TWA$TWA_5d,TWA$TWA_5d), col = "black", lwd = 2.5) # add TWA SDB disk exposed 5 days
lines(c(2.5,4.5),c(TWA$TWA_2d_7da,TWA$TWA_2d_7da), col = "red", lwd = 2.5) # add TWA t = 2 days based on SDB disk exposed for 7 days
lines(c(2.5,4.5),c(TWA$TWA_2d_5da,TWA$TWA_2d_5da), col = "orange", lwd = 2.5) # add TWA t = 2 days based on SDB disk exposed for 5 days

legend('topright', 
       legend = c("7 day disk, t = 2", "5 day disk, t = 2", "5 day disk, t = 5", "7 day disk, t = 7", "water concentration"),
       col = c("red", "orange", "black", "darkgreen", "blue"),
       lty = c(rep(1, 5)), bty = "n", cex=1, lwd = 2)
dev.off()
```


### prepare Figure S6 in the SI
```{r}

for(i in 1:length(Compounds)) { # number of compounds 
y <- Compounds[i]

  # only use the compound y for this run, runs once through each compound
  Water_t1 = Results_Water[Results_Water$Compound == y,] 
  TWA = TWA_SDB[TWA_SDB$Compound == y,]


# now plotting + saving each of the fittings to check if everything worked
file = y
file = file.path(path,"TWA", paste(file,"png", sep="."))

png(filename=file, width=7, height=6, units="in", res=1000)
par(mar=c(4.1, 4.9, 2.1, 2.1))
# create empty plot
plot(Water_t1$Time.water, Water_t1$Conc.water.D1, type ="n", pch = 16, ylim = c(0, (11/10)*max(Water_t1$Conc.water.D1)),
     xlim = c(0,7), main = paste(y), lwd = 2, ylab = expression(paste("concentration [ng L"^"-1"*"]")), xlab = c("time [days]"), las = 1, cex.axis = 1.6, cex.lab = 1.5, cex.main = 1.5, xaxs="i", yaxs="i")


# Water concentration
loess_water <- loess(Conc.water.D1  ~ Time.water, data = Water_t1, span = 0.3)
lines(Water_t1$Time.water, predict(loess_water), col = "blue", lwd = 2)
points(Water_t1$Time.water, Water_t1$Conc.water.D1, col = "blue", pch = 16)

abline(h = TWA$TWA_7d, col = "darkgreen", lwd = 2.5) # add TWA SDB disk exposed 7 days
lines(c(2,7.5),c(TWA$TWA_5d,TWA$TWA_5d), col = "black", lwd = 2.5) # add TWA SDB disk exposed 5 days
lines(c(2.5,4.5),c(TWA$TWA_2d_7da,TWA$TWA_2d_7da), col = "red", lwd = 2.5) # add TWA t = 2 days based on SDB disk exposed for 7 days
lines(c(2.5,4.5),c(TWA$TWA_2d_5da,TWA$TWA_2d_5da), col = "orange", lwd = 2.5) # add TWA t = 2 days based on SDB disk exposed for 5 days

dev.off()

}

```



## correlation previous sampling rates
create dataframe where sampling rates calculated in this study are added
```{r}
Rs_compare <- merge(Rs_other_studies[,c(2, 4, 6, 7, 11, 10)], Rs_prop[,c(1, 2, 3)], by = c("CAS"))
```

### without closing membrane == naked disks
```{r}
Rs_compare_naked <- Rs_compare[is.na(Rs_compare$Membran),]
Rs_compare_naked <- droplevels(Rs_compare_naked)
```


### Prepare Figure 2
```{r}
svg(filename="comparison_Rs.svg", width=8, height=6.5, pointsize=14)
par(mar=c(5.5, 6.5, 2.1, 2.1), mgp = c(4, 1, 0))


plot(Rs_compare_naked$Rs, Rs_compare_naked$Rs_comb, type="n", xlab = NA , ylab= NA, pch=16, ylim=c(0, 1.3), xlim=c(0, 1.3), cex.axis = 1.2, cex.lab = 1.2, las = 1, xaxs="i", yaxs="i")
mtext(side = 1, expression(atop("sampling rate [L d"^"-1"*"]", "previous studies")), line = 4, cex = 1.2)
mtext(side = 2, expression(atop("sampling rate [L d"^"-1"*"]", "this study")), line = 3, cex = 1.2)

cols <- c(NA, NA, "steelblue", NA, "black", "darkgreen", "red", NA, NA)
pchs <- c(15, 19, NA, NA, 17, NA, 18, NA, NA, NA, 7)
points(Rs_compare_naked$Rs, Rs_compare_naked$Rs_comb, col = cols[Rs_compare_naked$Flow.factor.new], # colour for flow velocity
       pch = pchs[Rs_compare_naked$publication.factor], cex=1.4) # shape for publication

abline(0,1, col = "red") # identity line, 1:1 line

# Problem: two compounds have the same first three letters, correct abbreviations
Rs_compare_naked$Comp_short <- substring(Rs_compare_naked$Compound, 1 , 3) 
Rs_compare_naked$Comp_short <- ifelse(Rs_compare_naked$Compound == "Dimethoat", "Dit", 
                                       ifelse(Rs_compare_naked$Compound == "Dimethomorph", "Dip",
                                              Rs_compare_naked$Comp_short))
text(Rs_compare_naked$Rs, Rs_compare_naked$Rs_comb, labels=Rs_compare_naked$Comp_short,  pos=3, offset=0.5,cex=0.8)

legend('bottomright', 
       legend = c("0.13 - 0.14 m s", "0.15 - 0.2 m s", "0.4 m s", "0.75 - 0.85 m s"),
       col = c("steelblue", "black", "darkgreen", "red"),
       pch = c(rep(16, 4)), bty = "n", cex=1)

legend('topright', 
       legend = expression(atop("s"^"-1")), bty = "n", cex=1)

dev.off()
```
saved as svg, overlaying compound names were moved using Inkscape.

#### measured divergence of correlation to the identity line
use Q2 statistics
```{r}
 # methods, stats
used_data <- Rs_compare_naked[, c("Rs_comb", "Rs")]
x <- c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0)
y <- c(0.1, 0.2, 0.3, 0.4, 0.5, 0.6, 0.7, 0.8, 0.9, 1.0)
data_1_1 <- as.data.frame(cbind(x,y))
names(data_1_1)[1] <- c("Rs_comb")
names(data_1_1)[2] <- c("Rs")

result <- q2(used_data, data_1_1, Rs_comb ~ Rs)

# calculate RMSE
result@result$pred$rmse
```
